---
title: "First Draft"
author: "Group 10"
date: "5/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Packages
library(ggplot2)
library(plotly)
library(dplyr)
library(tidyverse)
library(caret)
```

```{r}
# bring in the data
stroke = read_csv("stroke.csv")
```


# Question
•	Question and background information on the data and why you are asking this question(s).  References to previous research/evidence generally would be nice to include. – You must present your question to me during office hours, either next week on  26th or the following week on the 3rd  

# Exploratory Data Analysis
•	Exploratory Data Analysis – Initial summary statistics and graphs with an emphasis on variables you believe to be important for your analysis. 

# Methods and Evaluation
Do evaluation with each method, not all after
  
•	Methods – Techniques you are using to address your question and the results of those methods.
•	Evaluation of your model – Select appropriate metrics and explain the output as it relates to your question.

## KNN

## Clustering

## Decision

## Random Forest

# Fairness Assessment
•	Fairness assessment – if necessary, should you happen to have any protected classes.

# Conclusions
•	Conclusions – What can you say about the results of the methods section as it relates to your question given the limitations to your model. 

# Future Work
•	Future work – What additional analysis is needed or what limited your analysis on this project.
